Hydras \& Co.

### 2023-04-27 doc/chapter-fol.tex
First experiments on Fol notations
Examples of new notations (doc / MoreAck)

A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments).

@@

\subsubsection{A toy language}
In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
two constant symbols: $a$ and $b$, a function symbol $f$ (of arity 1), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.

@@

First, we declare two ranked alphabets. 
\inputsnippets{Languages/LNTDef1}

@@

@@

\end{remark}
\subsubsection{Examples}

With our toy language, we can build the following terms, respectively written $a$, $f(a)$ and $f(f(v_1))$ in usual mathematical notation.

Writing these terms in bare \gallina syntax would result in quite clumsy scripts. Thus, we may use a still experimental notation scope which allows us to write/print terms and formulas in a 
notation similar to \coq's.

We put both the associated \coq{} terms, in bare \gallina{} syntax, and with a still experimental and provisional notation, defined in
~\href{../theories/html/hydras.MoreAck.FOL_notations.html}{MoreAck.FOL\_notations}).

\inputsnippets{FolExamples/toyNotation}

\inputsnippets{FolExamples/smallTerms}

For instance the term $v_1+0$, where $v_1$ is a variable,
is represented by a term of type (\texttt{Term LNN}).

@@

Let us give a few examples of first-order formulas.

\label{fol:examplesf1f2f3}
\begin{enumerate}
\item $\forall\,v_0, v_0=0 \vee \exists\;v_1,\, v_0= S\;v_1$
\item $\exists\,v_2,\, 0< v_2 \wedge 4=v_2+v_2$
\item $ v_0=0 \vee \exists\;v_1,\, v_0= S\;v_1$
\item $v_0= v_1 + v_1 \leftrightarrow v_0 = v_1 \times 2$
\end{enumerate}
\inputsnippets{FolExamples/f1Example}

\begin{remark}[bound variables]
\inputsnippets{FolExamples/boundVars}  
\end{remark}

@@

>>>>

A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments). As suggested by Russel O'Connor in~\cite{OConnor05}, we consider two arity functions instead of a function defined on the sum type \texttt{Relations + Functions}.

@@

\subsubsection{A toy language}
In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
two constant symbols: $a$ and $b$, three function symbols $f$, $g$ and $h$  (of respective arity 1,1 and 2), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.

From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}.

@@

\begin{remark}
  The constructors of types \texttt{Rel} and \texttt{Fun} are suffixed by an underscore, in order to reserve the names \texttt{a}, \texttt{f}, \texttt{h}, \texttt{R}, etc. to functions which build terms and formulas
(please look at next sections).

\end{remark}

@@

\inputsnippets{FolExamples/TermExamples1}

@@

\end{remark}

\begin{remark}
 We can use \coq's \texttt{Notation} features to print and parse terms in a more readable form.

\inputsnippets{FolExamples/beforeAbbreviations, FolExamples/termAbbreviations, FolExamples/NowItsBetter}

\end{remark}

@@

Let us give a few examples of first-order formulas. on our toy language.

\label{fol:examplesf1f2f3}
\begin{enumerate}
\item $R\;a\;b$
\item $\forall v_0 v_1, (R\;v_0\;v_1 \leftrightarrow R\;v_1\;v_0$
\item $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
\end{enumerate}

\inputsnippets{FolExamples/FormExamples}

If we define a few new notations, the notation scope \texttt{fol\_scope} allow us to parse and print formulas in a less clumsy way.

From href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
\inputsnippets{fol/folScope}

\inputsnippets{FolExamples/toyNotationForm}

\inputsnippets{FolExamples/toyNotationForm2}
\begin{remark}[bound variables]
\inputsnippets{FolExamples/boundVars}  
\end{remark}

First, we declare two ranked alphabets.

\begin{remark}
  We depart a little from \cite{Goedel}'s notations, where the 
function and relation symbols are called \texttt{Plus}, 
\texttt{Mult}, \texttt{LT}, etc. In our version, these type constructors are called \texttt{Plus\_}, 
\texttt{Mult\_}, \texttt{LT\_}, etc., while the names without final underscores are bound to term building functions (\emph{e.g.}
the function which takes two terms and builds the term representing their sum).
\end{remark}

\inputsnippets{Languages/LNTDef1}

\begin{todo}
Start with toy Language , then LNT and LNN.
\end{todo}

% \subsubsection{Examples}

% With our toy language, we can build the following terms, respectively written $a$, $f(a)$ and $f(f(v_1))$ in usual mathematical notation.

% Writing these terms in bare \gallina syntax would result in quite clumsy scripts. Thus, we may use a still experimental notation scope which allows us to write/print terms and formulas in a 
% notation similar to \coq's.

% We put both the associated \coq{} terms, in bare \gallina{} syntax, and with a still experimental and provisional notation, defined in
% ~\href{../theories/html/hydras.Ackerm.FOL_notations.html}{MoreAck.FOL\_notations}).

% \inputsnippets{FolExamples/toyNotation}

% \inputsnippets{FolExamples/smallTerms}

For instance the term $v_1+0$, where $v_1$ is a variable,
is represented by the following \gallina term of type 
(\texttt{fol.Term LNT}).

@@

\begin{todo}
  Present the general issue about evaluation, and our provisional solution.
\end{todo}
